
WEEK 5
Feature Selection & Lasso
A fundamental machine learning task is to select amongst a set of features to include in a model. In this module, you will explore this idea in the context of multiple regression, and describe how such feature selection is important for both interpretability and efficiency of forming predictions.
To start, you will examine methods that search over an enumeration of models including different subsets of features. You will analyze both exhaustive search and greedy algorithms. Then, instead of an explicit enumeration, we turn to Lasso regression, which implicitly performs feature selection in a manner akin to ridge regression: A complex model is fit based on a measure of fit to the training data plus a measure of overfitting different than that used in ridge. This lasso method has had impact in numerous applied domains, and the ideas behind the method have fundamentally changed machine learning and statistics. You will also implement a coordinate descent algorithm for fitting a Lasso model.

Coordinate descent is another, general, optimization technique, which is useful in many areas of machine learning.

\item Slides presented in this module
\item The feature selection task
\item All subsets
\item Complexity of all subsets
\item Greedy algorithms
\item Complexity of the greedy forward stepwise algorithm
\item Can we use regularization for feature selection?
\item Thresholding ridge coefficients?
\item The lasso objective and its coefficient path
\item Visualizing the ridge cost
\item Visualizing the ridge solution
\item Visualizing the lasso cost and solution
\item Download the notebook and follow along
\item Lasso demo
\item What makes the lasso objective different
\item Coordinate descent
\item Normalizing features
\item Coordinate descent for least squares regression (normalized features)
\item Coordinate descent for lasso (normalized features)
\item Assessing convergence and other lasso solvers
\item Coordinate descent for lasso (unnormalized features)
\item Deriving the lasso coordinate descent update
\item Choosing the penalty strength and other practical issues with lasso
\item A brief recap
Quiz · Feature Selection and Lasso
\item Reading: Using LASSO to select features
Quiz · Using LASSO to select features
\item Reading: Implementing LASSO using coordinate descent
Quiz · Implementing LASSO using coordinate descent
