\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{fancyhdr}
%\usepackage{listings}
\usepackage{framed}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{fancyhdr}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2570}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{LastRevised=Wednesday, February 23, 2011 13:24:34}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

\pagestyle{fancy}
\setmarginsrb{20mm}{0mm}{20mm}{25mm}{12mm}{11mm}{0mm}{11mm}
\lhead{MA4128} \rhead{Mr. Kevin O'Brien}
\chead{Advanced Data Modeling}
%\input{tcilatex}

\begin{document}
\section*{Clustering Analysis}
\begin{enumerate}
\item What is the difference between agglomerative and divisive clustering?
\item What is the difference between hierarchical and non-hierarchical clustering?
\item How do we determine the appropriate number of clusters?  Give two different visualization methods that are used to display the outcome of a cluster analysis.

\item In the context of hierarchical cluster analyses, distinguish between agglomerative clustering and divisive clustering.
\item What is a vertical icicle plot used for? Give a brief dscription, supporting your answer
with sketches.

\item Explain why squared Euclidean Distance may be used in preference to Euclidean Distance.
\item Compute the squared Euclidean distance and the Manhattan (a.k.a. city block) distance between the following points A and B.
\[ \mbox{A} =\{2,5,1,7\}  \]
\[ \mbox{B} =\{1,6,3,6\}  \]
\item What is a dendrogram? What does does a dendrogram depict?% Question
\item What is the purpose of a cluster analysis? % Question

\item What is two-step clustering analysis?

\item Compare and contrast any two linkage methods. % Question

\item Explain the difference between Ward's method and K-means clustering.% Question
\item Discriminant Analysis is very similar to Clustering Analysis, but one key difference. Explain this difference.% Question
\item What is the difference between a linkage method and a similarity measure? Discuss the purpose of both.% Question
\item Compare and contrast any two linkage methods.
\item Discus how to determine the appropriate number of clusters.
\end{enumerate}
\end{document}
