\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{fancyhdr}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2570}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{LastRevised=Wednesday, February 23, 2011 13:24:34}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

\pagestyle{fancy}
\setmarginsrb{20mm}{0mm}{20mm}{25mm}{12mm}{11mm}{0mm}{11mm}
\lhead{MA4128} \rhead{Mr. Kevin O'Brien}
\chead{Advanced Data Modelling}
%\input{tcilatex}

\begin{document}
	\section*{Binary Logistic Regression: Tutorial Sheet}
\begin{enumerate}
	
	\item 
Under what sort of modelling problem would you use Binary Logistic Regression?

	\item What is a logit? How can you transform it into a probability?
	\item Suppose that the probability of a success is 0.70. Compute the corresponding odds.
	\item The usual assumptions placed on the error terms in ordinary least squares
	regression are:
	\begin{itemize}
		\item[$\ast$] independently distributed
		\item[$\ast$] identically distributed (equal variance)
		\item[$\ast$] normally distributed
	\end{itemize}
	Which of these assumptions are violated when dealing with binary response
	data? Explain briefly how each is violated.
	\item What is a dummy variable? Explain how it is used in Logistic Regression. Support your answer with an example.
	\item There are three variants for the forward selection procedures used by SPSS. Name these three.
	\item What is the Likelihood Ration Test? Describe how it used in Logistic regression.
	\item Suppose the odds of an outcome are 4. What is the probability of that outcome?
	\item Suppose the probability of an outcome is 70\%. What is the odds of that outcome occuring?
	\item What is a logit? how is it computed into a probability?
	\item Suppose that, out of a samle of 100 women and 100 men, 80 men drank alcohol in the last week, while 20 women drank alcohol in past week. Compute the odds ratio for Women to men
	\item What is logistic regression? How does it differ from linear regression? Under what circumstances would you use it?
\end{enumerate}	
\end{document}